---
title: "predicting Healthy Ride ridership with {prophet}"
author: "Conor Tompkins"
date: "6/21/2019"
output: html_document
---

This post is about predicting demand for the Healthy Ride bike system in Pittsburgh. I wanted to try out Facebook's `prophet` package and try to do some time series forecasting.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE)
```

As usual, load the required packages and set up the environment:
```{r}
library(tidyverse)
library(lubridate)
library(prophet)
library(janitor)
library(hrbrthemes)

options(scipen = 999)

theme_set(theme_bw())
```

This code loads the data and formats the date column so the `prophet` package can interface with it. I use `dir()` to find all the CSV files in  the folder and then `set_names()` and `map_df(read_csv())` to read each of the CSV files into memory.
```{r}
source("scripts/load_data.R")
```

The data I will use contains the number of rides per day and also includes the month and year. `prophet` will identify the time series patterns ("seasonality") in the data and identify the "true" pattern

`prophet` has a plug-and-play workflow that is easy to use, but it has more stringent requirements for how the data has to be shaped. The date data has to be named `ds` and the target variable has to be named `y`. 
```{r}
df <- df %>% 
  count(date) %>% 
  rename(ds = date,
         y = n)
```

```{r}
glimpse(df)
```

This code fits a model to the dataset.
```{r}
m <- prophet(df)
```

`make_future_dataframe()` creates the dataframe that `prophet` uses to make its forecast. In this case, I have it create a dataframe with 365 days of additional rows to predict onto.
```{r}
future <- make_future_dataframe(m, periods = 365, freq = "day")
```

This code performs the forecast on the future dataset.
```{r}
forecast <- predict(m, future) %>% 
  as_tibble()
```

The output is a dataframe with the date, the predicted ridership, and the upper and lower bounds of the prediction.
```{r}
forecast
```

`plot` automatically plots the `forecast` data:
```{r}
plot(m, forecast)
```

`prophet` also decomposes the various time series seasonalities:
```{r}
prophet_plot_components(m, forecast)
```