---
title: Clustering Healthy Ride Trips
author: Conor Tompkins
date: '2017-11-12'
slug: clustering-healthy-ride-trips
categories:
  - Pittsburgh
  - R
tags:
  - Healthy Ride
  - Pittsburgh
  - R Markdown
  - WPRDC
  - kmeans
draft: true
---

We can also apply kmeans clustering to the data to identify types of trips. This method only works with continuous data, so we will only be working with latitudes and longitudes.

First, get the top 50 stations in terms of trips:
```{r get_stations_for_kmeans}
df_long %>% 
  count(station_name, sort = TRUE) %>% 
  top_n(50) %>% 
  select(station_name) %>% 
  unlist() -> top_stations
```

Then, remove all the rows with NA values and scale the data:
```{r munge_data_for_kmeans}
df_wide <- df_wide %>%
  na.omit() %>% 
  select(from_longitude, from_latitude, to_longitude, to_latitude) %>% 
  mutate_all(scale)
```

Perform the clustering algorithm:
```{r perform_kmeans}
kclust <- kmeans(df_wide, 7)
kclust
summary(kclust)
```

Attempt various numbers of clusters (from 1 to 7) to find the best fit (warning, this is memory and compute intensive)
```{r various_clusters}
kclusts <- data.frame(k=1:7) %>% group_by(k) %>% do(kclust=kmeans(df_wide, .$k))
```

Pull out the cluster data and tidy it:
```{r tidy_clusters}
clusters <- kclusts %>% group_by(k) %>% do(tidy(.$kclust[[1]])) %>% 
  rename(from_longitude = x1,
         from_latitude= x2,
         to_longitude = x3,
         to_latitude = x4)
#assignments <- kclusts %>% group_by(k) %>% do(augment(.$kclust[[1]], df_wide))
clusterings <- kclusts %>% group_by(k) %>% do(glance(.$kclust[[1]]))
```

This plot tells us that 6 clusters is probably the correct number. The goal is to find the lowest tot.withinss value with the fewest number of clusters:
```{r scree}
ggplot(clusterings, aes(k, tot.withinss)) + 
  geom_line() +
  scale_x_continuous(breaks = 1:10)
```

Since we scaled the data, it no longer has meaning when applied to a map. However, we can see the general types of trips:
```{r plot_kmeans}
clusters %>% 
  filter(k >= 2) %>% 
  ggplot(aes(x = from_longitude, xend = to_longitude, y = from_latitude, yend = to_latitude)) + 
  geom_segment(aes(color = cluster, size = size),
               arrow = arrow(length = unit(0.03, "npc"))) + 
  facet_wrap(~ paste(k, "clusters"))
```